---
title: "experiment"
format: html
---

# Libraries and Data

## Libraries

```{r}
#| label: libraries

library(coda)
library(MCMCpack)
library(BayesVarSel)
library(BMS)
library(BoomSpikeSlab)
library(spikeSlabGAM)
library(bayeslm)
library(BayesSubsets)
#library(EMVS)  # Man
library(tidyverse)
```


## Data

Here is a table that describes each model and its characteristics:

| Name | n | p | p_sig |  n vs. p | Description |
|------|---|---|-------|----------|-------------|
| Model 1 | 1000 | 10 | 3 | n >> p | Few predictors, sparse |
| Model 2 | 1000 | 950 | 750 | n > p | n slightly > p, dense |
| Model 3 | 1000 | 950 | 5 | n > p | n slightly > p, sparse |
| Model 4 | 1000 | 1050 | 950 | n < p | n slightly < p, dense |
| Model 5 | 1000 | 1050 | 5 | n < p | n slightly < p, sparse |
| Model 6 | 1000 | 10000 | 9500 | Many predictors, dense |
| Model 7 | 1000 | 10000 | 5 | Many predictors, sparse

```{r}
#| label: generate data

set.seed(465)

model_1 <- simulate_lm(1000, 10, p_sig = 3)
model_2 <- simulate_lm(1000, 950, p_sig = 750)
model_3 <- simulate_lm(1000, 950, p_sig = 5)
model_4 <- simulate_lm(1000, 1050, p_sig = 950)
model_5 <- simulate_lm(1000, 1050, p_sig = 5)
model_6 <- simulate_lm(1000, 10000, p_sig = 9500)
model_7 <- simulate_lm(1000, 10000, p_sig = 5)
```

## Hyperparameters

```{r}
#| label: defining hyperparameters

N_SAMPLES <- 30000
BURN_IN <- 5000
```



# Non-Prior Methods

## Subset Selection

```{r}
#| label: run bayeslm and evaluate

set.seed(465)

X <- model_1$X
y <- model_1$y

model_1_fit <- bayeslm(y ~ X[,-1],
                       N = N_SAMPLES,
                       burnin = BURN_IN,
                       prior = "ridge"
                       )

temp <- post_predict(post_y_hat = tcrossprod(model_1_fit$beta, X),
                    post_sigma = model_1_fit$sigma,
                    yy = y)
post_y_pred = temp$post_y_pred
post_lpd = temp$post_lpd

indicators_1 <- branch_and_bound(yy = fitted(model_1_fit),
                              XX = X)

accept_info <- accept_family(post_y_pred = post_y_pred,
                            post_lpd = post_lpd,
                            XX = X,
                            indicators = indicators_1,
                            yy = y,
                            post_y_hat = tcrossprod(model_1_fit$beta, X))



# How many subsets are in the acceptable family?
length(accept_info$all_accept) # 76

# These are the rows of `indicators` that belong to the acceptable family:
head(accept_info$all_accept)

# An example acceptable subset:
ex_accept <- accept_info$all_accept[1]
which(indicators[ex_accept,])

# Simplest acceptable subset:
beta_hat_small <- accept_info$beta_hat_small

# Which coefficients are nonzero:
S_small <- which(beta_hat_small != 0)

# How many coefficients are nonzero:
length(S_small) # 4

# Acceptable subset that minimizes CV error:
beta_hat_min <- accept_info$beta_hat_min

# Typically much larger (and often too large...)
sum(beta_hat_min != 0) # 5

# Variable importance metrics
vi_e <- var_imp(indicators = indicators_1,
               all_accept = accept_info$all_accept)$vi_inc

# Variables appearing in all acceptable subsets
all_acceptable <- which(vi_e == 1) # Correct variables

# Size
length(all_acceptable) # 4
```



